### 2023-08-18 header.tex
Upload all files

% **************GENERATED FILE, DO NOT EDIT**************

\title{Pigeons.jl: Distributed sampling from intractable distributions}

\author[1]{Nikola Surjanovic}
\author[1]{Miguel Biron-Lattes}
\author[2]{Paul Tiede}
\author[3]{Saifuddin Syed}
\author[1]{Trevor Campbell}
\author[1]{Alexandre Bouchard-Côté}
\affil[1]{University of British Columbia}
\affil[2]{Harvard University}
\affil[3]{University of Oxford}

\keywords{Distributed computation, Bayesian inference, parallelism invariance, 
multi-threading, Message Passing Interface, Markov chain Monte Carlo}

\hypersetup{
pdftitle = {My JuliaCon proceeding},
pdfsubject = {JuliaCon 2019 Proceedings},
pdfauthor = {1st author, 2nd author, 3rd author},
pdfkeywords = {Julia, Optimization, Game theory, Compiler},
}