### 2023-10-31 paper/conclusion.tex
use paper folder

\section{Conclusion}
Pigeons.jl is a Julia package that enables users with no experience in distributed 
computing to efficiently approximate posterior distributions and solve challenging 
Lebesgue integration problems over a distributed computing environment.
The core algorithm behind Pigeons.jl is distributed, non-reversible parallel tempering 
\cite{syed2021nrpt,surjanovic2022vpt}. Pigeons.jl can be used in a multi-threaded context, 
as well as distributed over up to thousands of MPI-communicating machines. 
Further, Pigeons.jl is designed so that for a given seed, the output is \textit{identical}
regardless of the number of threads or processes used.