### 2025-05-06 workshop/paper.tex

\documentclass{article}

\usepackage[accepted]{icml2025}
\usepackage{listings}
\usepackage{jlcode}
\usepackage{natbib}
\setcounter{page}{1}

\usepackage{expl3} % add missing functionality in the old texlive version running on the server 
\usepackage[autonum, blackhypersetup]{shortex}
\usepackage{xcolor}
%\let\subcaption\undefined % overwrite 'subcaption' from the style file
%\usepackage[font=scriptsize]{subcaption}
\usepackage{forest}

\newcommand{\NS}[1]{\textcolor{red}{#1}}
\def\bfx{\mathbf{x}}
\def\bfX{\mathbf{X}}

\definecolor{julia1}{rgb}{0, 0.60, 0.98}
\definecolor{julia2}{rgb}{0.89, 0.43, 0.28}
\definecolor{julia3}{rgb}{0.24, 0.64, 0.30}
\definecolor{julia4}{rgb}{0.76, 0.44, 0.82}
\definecolor{julia5}{rgb}{0.67, 0.55, 0.09}
\definecolor{julia6}{HTML}{00aaae}
\definecolor{julia7}{HTML}{ed5e93}
\definecolor{julia8}{HTML}{c68225}

\pdfsuppresswarningpagegroup=1 % avoid warning due to pdfs made with photoshop

% The \icmltitle you define below is probably too long as a header.
% Therefore, a short form for the running title is supplied here:
\icmltitlerunning{Pigeons.jl: Distributed sampling from intractable distributions}

\begin{document}

\twocolumn[
\icmltitle{Pigeons.jl: Distributed sampling from intractable distributions}

\icmlsetsymbol{equal}{*}

\begin{icmlauthorlist}
\icmlauthor{Nikola Surjanovic}{ubc}
\icmlauthor{Miguel Biron-Lattes}{sfu}
\icmlauthor{Paul Tiede}{harvard}
\icmlauthor{Saifuddin Syed}{oxford}
\icmlauthor{Alexandre Bouchard-C{\^o}t{\'e}}{ubc}
\icmlauthor{Trevor Campbell}{ubc}
\end{icmlauthorlist}

\icmlaffiliation{harvard}{Center for Astrophysics, Harvard University, United States}
\icmlaffiliation{oxford}{Department of Statistics, University of Oxford, United Kingdom}
\icmlaffiliation{sfu}{Department of Statistics and Actuarial Sciences, Simon Fraser University, Canada}
\icmlaffiliation{ubc}{Department of Statistics, University of British Columbia, Canada}

\icmlcorrespondingauthor{Miguel Biron-Lattes}{mbironla@sfu.ca}
\icmlkeywords{Distributed computation, Bayesian inference, Markov chain Monte Carlo, Parallelism Invariance, Message Passing Interface, Multi-Threading}

\vskip 0.3in
]

\printAffiliationsAndNotice{}  % leave blank if no need to mention equal contribution

\input{abstract}
\input{introduction.tex}
\input{examples.tex}
\input{PT.tex}
\input{SPI.tex}
\input{related.tex}
\input{conclusion.tex}
\input{ack}

\bibliographystyle{icml2025}
\bibliography{../paper/ref.bib}

\end{document}

% Inspired by the International Journal of Computer Applications template