### 2025-05-07 workshop/abstract.tex

\begin{abstract}
Sampling from complicated probability distributions---such as 
multimodal posteriors arising in Bayesian inference and high-dimensional 
distributions in statistical mechanics---is a challenging problem. Pigeons.jl provides 
a simple user interface to perform such computations single-threaded, multi-threaded,
and/or distributed over thousands of MPI-communicating machines. In this paper,
we discuss a feature of Pigeons.jl that we call \emph{strong parallelism invariance} (SPI): 
a guarantee that the output for a given seed is identical irrespective of the number 
of threads and processes. This feature is crucial for scientific reproducibility and 
software validation. We describe the key features of Pigeons.jl that enable a 
distributed and randomized algorithm satisfying SPI.
\end{abstract}