### 2025-05-07 workshop/paper.tex

\documentclass{article}

\input{macros}

% The \icmltitle you define below is probably too long as a header.
% Therefore, a short form for the running title is supplied here:
\newcommand{\thetitle}{Reproducible sampling from intractable distributions with Pigeons.jl}
\icmltitlerunning{\thetitle}

\begin{document}

\twocolumn[
\icmltitle{\thetitle}

\icmlsetsymbol{equal}{*}

\begin{icmlauthorlist}
\icmlauthor{Nikola Surjanovic}{ubc}
\icmlauthor{Miguel Biron-Lattes}{sfu}
\icmlauthor{Paul Tiede}{harvard}
\icmlauthor{Saifuddin Syed}{oxford}
\icmlauthor{Alexandre Bouchard-C{\^o}t{\'e}}{ubc}
\icmlauthor{Trevor Campbell}{ubc}
\end{icmlauthorlist}

\icmlaffiliation{harvard}{Center for Astrophysics, Harvard University, United States}
\icmlaffiliation{oxford}{Department of Statistics, University of Oxford, United Kingdom}
\icmlaffiliation{sfu}{Department of Statistics and Actuarial Sciences, Simon Fraser University, Canada}
\icmlaffiliation{ubc}{Department of Statistics, University of British Columbia, Canada}

\icmlcorrespondingauthor{Miguel Biron-Lattes}{mbironla@sfu.ca}
\icmlkeywords{Distributed computation, Bayesian inference, Markov chain Monte Carlo, Parallelism Invariance, Message Passing Interface, Multi-Threading}

\vskip 0.3in
]

\printAffiliationsAndNotice{}  % leave blank if no need to mention equal contribution

\input{abstract}
\input{introduction.tex}
\input{SPI.tex}
\input{related.tex}
%\input{conclusion.tex}
%\input{ack}

\bibliographystyle{icml2025}
\bibliography{../paper/ref.bib}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% APPENDIX
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\appendix
\onecolumn

\input{examples}
\input{PT}

\end{document}

% Inspired by the International Journal of Computer Applications template