### 2025-05-13 workshop/new_developments.tex
new devels + fixes

\section{Recent developments}

\paragraph{autoMALA}
Since version v0.2.0, Pigeons.jl's default explorer for most models in continuous
space has been autoMALA \citep{biron2024automala}, a novel MCMC sampler
based on the Metropolis-adjusted Langevin algorithm (MALA). autoMALA continually
adapts its step size at each iteration based on the local geometry of the target
distribution. Such automatic adjustments are highly desirable in the local exploration 
phase of PT (see \cref{sec:PT} for a review of PT), since otherwise we would 
need to find optimal step sizes for each distribution in the annealing path---a 
prohibitively costly process.

\vspace{-1em}

\paragraph{Support for BUGS models}
The latest version of Pigeons.jl at the time of writing (v0.4.9) added support 
for sampling from programs written in BUGS \cite{lunn2000winbugs,lunn2009bugs,
lunn2013bugs}, by leveraging the Julia implementation of the 
language in JuliaBUGS.jl\footnote{\url{https://turinglang.org/JuliaBUGS.jl/stable/}}.

\vspace{-1em}

\paragraph{Tempered Particle Gibbs for Turing-complete PPLs}
Particle Gibbs (PG, \citealp{andrieu2010particle}) is an MCMC
algorithm that has been shown \citep{wood2014new} to be a natural sampler for
arbitrary programs written in Turing-complete probabilistic programming 
languages (PPLs). For example, the Julia package Turing.jl \citep{ge2018turing}
provides a PG sampler that can handle any program written in DynamicPPL
\citep{tarek2020dynamicppl}, its underlying PPL.
However, this generality comes at the expense of poor performance in moderately
complex models. The reason is that PG is, essentially, a smart 
way for selecting samples from the prior that are close to the posterior.
When these distributions are very different, PG can be inefficient.

With Pigeons.jl we can partially alleviate this issue by using PG as an explorer
in PT. Indeed, PG should be efficient for exploring distributions closer to the 
prior. Such samples can then be transported towards the posterior chain via the PT process.
In order to construct the annealing sequence of distributions for an arbitrary 
PPL program, it suffices to inject an inverse temperature parameter 
$\beta\in[0,1]$ in the call to compute the log density of every \texttt{observe} 
statement \citep[see][for an explanation of this syntax]{gordon2014probabilistic} 
in the program. Leveraging multiple dispatch in Julia, we have
successfully applied this technique to DynamicPPL programs in our package
TPGExplorers.jl\footnote{\url{https://github.com/Julia-Tempering/TPGExplorers.jl}}.