### 2025-05-26 workshop/abstract.tex
NS commnets

\begin{abstract}
Sampling from complicated probability distributions---such as 
multimodal posteriors arising in Bayesian inference and high-dimensional 
distributions in statistical mechanics---is a challenging problem. \emph{Pigeons.jl} 
is a Julia package that implements \emph{Parallel Tempering}, a technique that 
leverages parallelism to improve the effectiveness of \emph{Markov chain Monte Carlo} 
algorithms. Pigeons.jl provides a simple user interface to perform such computations 
using single-threaded, multi-threaded, and/or distributed over thousands of MPI-communicating 
machines. In this paper, we discuss a feature of Pigeons.jl that we call \emph{strong 
parallelism invariance} (SPI): a guarantee that the output for a given seed is identical 
irrespective of the number of threads and processes. This feature is crucial for scientific 
reproducibility and software validation. We describe the key features of Pigeons.jl that 
enable a distributed and randomized algorithm satisfying SPI. Finally, we briefly discuss 
some recent developments in Pigeons.jl.
\end{abstract}